---
title: "theses.fr"
author: "Mathilde Mousset"
date: "14 décembre 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(hrbrthemes)
library(lubridate)
```



# General information

I downloaded all thesis defended in France and recorded on the <these.fr>[these.fr] datatabase on the 14 December 2018. This datafile does not take into account thesis that were never defended, or are still under completion to this date.

I first did some benign cleaning (remove non data lines, repeated headers, convert dates into universal format etc.) and saved the clean data.


```{r load_data}
# Load data
thesis <- read_csv2("./2_clean_data/thesis.csv")
```


There are PhDs with several authors, separated by commas. I filter them out.
Then, we fix the advisor columns. Right now, the advisor 1 is XY and the advisor 2 is YX. When there are two (or more) advisors, their names are in the same column, separated by a comma.


```{r}
thesis_clean <- thesis %>% 
  filter(!duplicated(thesis_id)) %>% 
  filter(!str_detect(author, pattern = ",")) %>% 
  select(-thesis_advisor2) %>% 
  mutate(advisors = str_split(thesis_advisor1, ","))
```



# Some general numbers

The number of thesis defended in France has decreased around the 2000, but rose again to similar level as before in the past decade.

```{r}
thesis_clean %>% 
  ggplot(aes(x = year_defense)) +
  geom_histogram(binwidth = 1) +
  labs(title = "Number of PhD defended in France since 1986",
       subtitle = "Why is there such a gap in the 2000 ?",
       caption = "Source: theses.fr",
       x = "Year of PhD defense",
       y = "Number of thesis") +
  theme_ipsum_rc()
```


Most PhDs have one or two supervisors. A couple have more.

```{r}
thesis_clean %>% 
  mutate(n_authors = lengths(advisors)) %>%
  count(n_authors) %>% 
  mutate(proportion = 100 * n / sum(n))
```


There are about 2500 persons with two PhDs (based on author ID), whith is probably correct. Whether the people with four to six PhD represent a problem in author ID attribution, real PhDs, or crackpots who managed to get a varnish of science is hard to tell. 

```{r}
thesis_clean %>% 
  # Remove authors with no ID
  drop_na(author_id) %>% 
  filter(author_id != ",") %>% 
  count(author_id, sort = TRUE) %>% 
  filter(n > 1) %>% 
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 1) +
  coord_flip() +
  labs(title = "Nb of people with more than one PhD",
       subtitle = "",
       x = "Number of PhDs",
       y = "Number of people",
       caption = "Source: theses.fr") +
  theme_ipsum_rc()
```


# Number of thesis per supervisor

While most people supervised or helped supervise a dozen of PhDs, a decent number of people supervised more than 20 PhDs, which I think as an awesome lot. In France, this is likely to depend on the field. In science doctoral schooln the number of PhD that a PI can supervise at any given time can be limited. I believe that it is a different culture in the humanities.


```{r}
thesis_clean %>% 
  drop_na(thesis_advisors_id) %>% 
  filter(!thesis_advisors_id == ",") %>% 
  separate_rows(thesis_advisors_id, sep = ",") %>% 
  drop_na(thesis_advisors_id) %>% 
  group_by(thesis_advisors_id) %>% 
  count(thesis_advisors_id, sort = TRUE) %>% 
  filter(n < 500) %>% 
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 0.1) +
  coord_flip() +
   scale_x_continuous(trans = "log10") +
     scale_y_continuous(trans = "log10") +
  labs(title = "Number of PhD supervised by advisor",
       subtitle = "Log10 scale",
       x = "Number of PhDs",
       y = "Number of advisor (or co-advisor)",
       caption = "Source: theses.fr") +
  theme_ipsum_rc()
```



# Questions

Questions :

- how many thesis per discipline
- time series

