---
title: "theses.fr"
author: "Mathilde Mousset"
date: "14 décembre 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(gganimate)
```

```{r my_theme, echo=FALSE}
plot_background_colour <- "#C2B3B4"
strip_background_colour <- "#C2B3B4"
grid_colour <- "#C2B3B4"

text_colour  <- "#3D1A1C"
title_colour <- "#3D1A1C"
axis_colour  <- "#3D1A1C"

bar_col <- "#92070E"

my_theme <- theme(
    # Text in general
    text = element_text(family = "Roboto Condensed",
                        colour = text_colour),
    
    # Title, subtitle, caption
    plot.title    = element_text(size = 16,
                                 lineheight = 1.6,
                                 face = "bold",
                                 colour = title_colour),
    plot.subtitle = element_text(size = 14,
                                 lineheight = 1),
    plot.caption  = element_text(size = 14),
    
    # Facet
    strip.background = element_rect(fill = strip_background_colour),
    strip.text = element_text(size = 12,
                              face = "bold"),

    
    # Axes
    axis.text  = element_text(size = 12),
    #axis.text.x = element_text(vjust = 1),
    #axis.text.y = element_text(hjust = 1),
    
    axis.title = element_text(size = 14,
                              face = "bold"),
    axis.line  = element_line(size = 0.5,
                              colour = axis_colour),

    # Legend
    legend.title = element_text(size = 12, 
                                face = "bold"), 
    legend.text = element_text(size = 12),
    legend.background = element_rect(fill = plot_background_colour),
    
    # Background
    panel.background = element_rect(fill = plot_background_colour),
    plot.background  = element_rect(fill = plot_background_colour),
    
    panel.grid.major = element_line(colour = grid_colour,
                                    size = 0.2,
                                    linetype = "dashed"),
    panel.grid.minor = element_line(colour = grid_colour, 
                                    size = 0.2, 
                                    linetype = "dashed"),
    panel.border = element_blank())
```


# General information

I downloaded all thesis defended in France and recorded on the <these.fr>[these.fr] datatabase on the 14 December 2018. This datafile does not take into account thesis that were never defended, or are still under completion to this date.

I first did some benign cleaning (remove non data lines, repeated headers, convert dates into universal format etc.) and saved the clean data.


```{r load_data}
load("./2_clean_data/thesis.RData")
```


There are PhDs with several authors, separated by commas. I filter them out.
Then, we fix the advisor columns. Right now, the advisor 1 is XY and the advisor 2 is YX. When there are two (or more) advisors, their names are in the same column, separated by a comma.


```{r}
thesis_clean <- thesis2 %>% 
  filter(!duplicated(thesis_id)) %>% 
  filter(!str_detect(author, pattern = ",")) %>% 
  select(-thesis_advisor2) %>% 
  mutate(advisors = str_split(thesis_advisor1, ","))



# Get months in English
thesis_clean$month_english = factor(thesis_clean$month_defense,
                                    ordered = TRUE)

levels(thesis_clean$month_english) <- c("January", "February", "March", "April", "May", "June", "July", "August", "September",  "October", "November", "December")
```



# Some general numbers

The number of thesis defended in France has decreased around the 2000, but rose again to similar level as before in the past decade.

```{r}
thesis_clean %>% 
  ggplot(aes(x = year_defense)) +
  geom_histogram(binwidth = 1,
                 fill = bar_col) +
  labs(title = "Number of PhD defended in France since 1986",
       subtitle = "Why is there such a gap in the 2000 ?",
       caption = "Source: theses.fr",
       x = "Year of PhD defense",
       y = "Number of thesis") +
  my_theme
```


Most PhDs have one or two supervisors. A couple have more.

```{r}
thesis_clean %>% 
  mutate(n_authors = lengths(advisors)) %>%
  count(n_authors) %>% 
  mutate(proportion = 100 * n / sum(n))
```


There are about 2500 persons with two PhDs (based on author ID), whith is probably correct. Whether the people with four to six PhD represent a problem in author ID attribution, real PhDs, or crackpots who managed to get a varnish of science is hard to tell. 

```{r}
thesis_clean %>% 
  # Remove authors with no ID
  drop_na(author_id) %>% 
  filter(author_id != ",",
         !author_id == "") %>% 
  
  count(author_id, sort = TRUE) %>% 
  filter(n > 1) %>% 
  
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 1,
                 fill = bar_col) +
  coord_flip() +
  labs(title = "People with more than one PhD",
       subtitle = "Some people hold more than three PhDs... O_O",
       x = "Number of PhDs",
       y = "Number of people",
       caption = "Source: theses.fr") +
  scale_y_continuous(breaks = seq(0, 3000, 200)) +
  my_theme
```


# Number of thesis per supervisor

While most people supervised or helped supervise a dozen of PhDs, a decent number of people supervised more than 20 PhDs, which I think as an awesome lot. In France, this is likely to depend on the field. In science doctoral schooln the number of PhD that a PI can supervise at any given time can be limited. I believe that it is a different culture in the humanities.


```{r}
thesis_clean %>% 
  drop_na(thesis_advisors_id) %>% 
  filter(!thesis_advisors_id == ",") %>% 
  separate_rows(thesis_advisors_id, sep = ",") %>% 
  drop_na(thesis_advisors_id) %>% 
  
  group_by(thesis_advisors_id) %>% 
  count(thesis_advisors_id, sort = TRUE) %>% 
  filter(n < 500) %>% 
  
  ggplot(aes(x = n)) +
  geom_histogram(binwidth = 0.1,
                 fill = bar_col) +
  coord_flip() +
   scale_x_continuous(trans = "log10") +
     scale_y_continuous(trans = "log10") +
  labs(title = "Number of PhD supervised by advisor",
       subtitle = "Log10 scale",
       x = "Number of PhDs per (co-)advisor",
       y = "Number of advisors",
       caption = "Source: theses.fr") +
  my_theme
```


# When are PhD defended?

```{r}
thesis_clean %>% 
  filter(year_defense >= 1990) %>% 
    filter(year_defense < 2018) %>% 
  group_by(year_defense) %>% 
  
  ggplot(aes(x = year_defense)) +
  geom_histogram(binwidth = 1,
                 fill = bar_col) +
  labs(title = "Number of PhD defended in France",
       subtitle = "From 1990 to 2017",
       x = "Year of PhD defense",
       y = "Number of PhD",
       caption = "Source: theses.fr") +
  my_theme
```






```{r}
thesis_clean %>% 
  filter(year_defense >= 2010) %>% 
    filter(year_defense < 2018) %>% 
  mutate(day_defense2 = day(date_defense)) %>% 
  group_by(year_defense,
           month_defense,
           day_defense2) %>% 
  count() %>% 
  
  ggplot(aes(x = month_defense, 
             y = day_defense2, 
             fill = n)) +
  facet_wrap(~year_defense, ncol = 8) +
  geom_tile() +
  
  labs(title = "Distribution of PhD defense over time",
       subtitle = "",
       x = "Day of the week",
       y = "Defense month") +
  # scale_fill_gradientn(colours = wes_palette("Zissou1", 
  #                                            100,
  #                                            type = "continuous")) +
  
  scale_fill_gradient(high = bar_col, 
                      low = title_colour) +
  
  my_theme +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.3))
```

Since the first of january  :

```{r}
thesis_clean %>% 
  filter(year_defense >= 2010) %>% 
    filter(year_defense < 2018) %>% 
    mutate(day_defense2 = day(date_defense)) %>% 
  filter(day_defense2 > 1) %>% 
  group_by(year_defense,
           month_defense,
           day_defense2) %>% 
  count() %>% 
  ggplot(aes(x = month_defense, 
             y = day_defense2, 
             fill = n)) +
  facet_wrap(~year_defense, ncol = 8) +
  geom_tile() +
  
  labs(title = "Distribution of PhD defense over time",
       subtitle = "",
       x = "Day of the week",
       y = "Defense month") +
  scale_fill_gradient(high = bar_col, 
                      low = title_colour) +
  my_theme +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.3))
```

```{r}
thesis_clean %>% 
  filter(year_defense >= 2010) %>% 
    filter(year_defense < 2018) %>% 
  mutate(day_defense2 = day(date_defense)) %>% 
  filter(day_defense2 > 1) %>% 
  group_by(year_defense,
           month_defense,
           day_defense) %>% 
  count() %>% 
  ggplot(aes(x = day_defense, 
             y = month_defense, 
             fill = n)) +
  facet_wrap(~year_defense, ncol = 8) +
  geom_tile() +
  
  labs(title = "Distribution of PhD defense over time",
       subtitle = "",
       x = "Day of the week",
       y = "Defense month") +
  
    scale_fill_gradient(high = bar_col, 
                      low = title_colour) +
 my_theme +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.3))
```



```{r}
thesis_clean %>% 
  filter(year_defense >= 2015,
         year_defense < 2018) %>% 
    mutate(year_month = format(as.Date(date_defense),
                              "%Y-%m")) %>% 
  group_by(year_defense, month_defense, year_month) %>% 
  count() %>% 
  ggplot(aes(x = month_defense,
             y = n,
             group = year_defense)) +
  geom_line(colour = bar_col,
            size = 2,
            alpha = 0.7) +
  annotate("text",
           x = "août",
                y = 2000,
                label = "Summer holidays",
                size = 5,
                colour = title_colour) +
  annotate("segment", 
           x = "août", 
           xend = "août", 
           y = 1800, 
           yend = 800, 
           colour = title_colour, 
           size = 1,
           arrow = arrow()) +
  
  labs(title = "Typical year calendar for PhD defenses",
       subtitle = "Most PhDs begin in autumn, and they have until December three years\nlater to defend, hence the end-of-year peak",
       x = "Month of defense",
       y = "Number of PhDs",
       caption = "Source: theses.fr") +
  my_theme
```

## Try gganimate

```{r}
p <- thesis_clean %>% 
  filter(year_defense < 2018,
         year_defense > 1989) %>% 
  group_by(year_defense) %>% 
  count() %>% 
  ggplot(aes(x = year_defense,
             y = n,
             group = 1)) +
  geom_line(size = 3, 
            colour = bar_col) +
  # geom_point(size = 3,
  #            colour = bar_col) +
  
  labs(title = "Number of PhD in France per year",
       subtitle = "Such a reduction in 2001",
       caption = "Source: theses.fr",
       x = "Year of defense",
       y = "Number of defended PhDs") +
  ylim(1, 14000) +
  my_theme +

  transition_reveal(year_defense) + 
  ease_aes("linear")

animate(p, 
        nframes = 100, 
        fps = 10)
```



# Ecology and evolution

## Get the data

```{r}
thesis_evology <- thesis_clean %>% 
  filter(str_detect(discipline, 
                    pattern = "Évolut|Evolut|evolut|évolut|Genetique des populations|Ecolog|ecolog|écolog|Écolog") |
           discipline %in% c("Physiologie et biologie des organismes et populations", "Biologie des populations et ecologie"))  %>% 
  drop_na(date_defense) %>% 
  filter(!str_detect(discipline,
                   pattern = "Terre|planete|lithos|geologie|Geologie|evolutionnaire|Materiaux|materiaux|gyneco|Gyneco"))
```


Explore this

```{r}
thesis_evology %>% 
  count(discipline, sort = TRUE) %>% View
```



## Timeline

```{r}
thesis_evology %>% 
  filter(year_defense >= 1990) %>% 
    filter(year_defense < 2018) %>% 
  group_by(year_defense) %>% 
  ggplot(aes(x = year_defense)) +
  geom_histogram(binwidth = 1,
                 fill = bar_col) +
  labs(title = "Number of PhD defended in ecology and evolution in France",
       subtitle = "From 1990 to 2017",
       x = "Year of PhD defense",
       y = "Number of PhD",
       caption = "Source: theses.fr") +
  my_theme
```



```{r}
thesis_evology %>% 
  filter(year_defense >= 2010) %>% 
    filter(year_defense < 2018) %>% 
  mutate(day_defense2 = day(date_defense)) %>% 
  group_by(year_defense,
           month_defense,
           day_defense2) %>% 
  count() %>% 
  ggplot(aes(x = month_defense, 
             y = day_defense2, 
             fill = n)) +
  facet_wrap(~year_defense, ncol = 8) +
  geom_tile() +
  
  labs(title = "Distribution of PhD defense over time",
       subtitle = "",
       x = "Day of the week",
       y = "Defense month") +
  scale_fill_gradient(high = bar_col, 
                      low = title_colour) +
  my_theme +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.3))
```


```{r}
thesis_evology %>% 
  filter(year_defense == 2010,
         month_defense == "janvier") %>% 
  head()
```


Since the first of january 
```{r}
thesis_evology %>% 
  filter(year_defense >= 2010) %>% 
    filter(year_defense < 2018) %>% 
    mutate(day_defense2 = day(date_defense)) %>% 
  filter(day_defense2 > 1) %>% 
  group_by(year_defense,
           month_defense,
           day_defense2) %>% 
  count() %>% 
  ggplot(aes(x = month_defense, 
             y = day_defense2, 
             fill = n)) +
  facet_wrap(~year_defense, ncol = 8) +
  geom_tile() +
  
  labs(title = "Distribution of PhD defense over time",
       subtitle = "",
       x = "Day of the week",
       y = "Defense month") +
  
scale_fill_gradient(high = bar_col, 
                      low = title_colour) +
  
  my_theme +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.3))
```

```{r}
thesis_evology %>% 
  filter(year_defense >= 2010) %>% 
    filter(year_defense < 2018) %>% 
  mutate(day_defense2 = day(date_defense)) %>% 
  filter(day_defense2 > 1) %>% 
  group_by(year_defense,
           month_defense,
           day_defense) %>% 
  count() %>% 
  ggplot(aes(x = day_defense, 
             y = month_defense, 
             fill = n)) +
  facet_wrap(~year_defense, ncol = 8) +
  geom_tile() +
  
  labs(title = "Distribution of PhD defense over time",
       subtitle = "",
       x = "Day of the week",
       y = "Defense month") +
  
  scale_fill_gradient(high = bar_col, 
                      low = title_colour) +
 my_theme +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.3))
```


```{r}
thesis_evology %>% 
  filter(year_defense == 2013,
         day_defense == "dimanche") %>% 
  view
```

Nothing scandalous.
