---
title: "Network analysis"
author: "Mathilde Mousset"
date: "22 février 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest) # To scrap PhD pages to get committe members

library(tidyverse)
#library(scales) 
library(hrbrthemes)
library(data.table)
# devtools::install_github("privefl/bigreadr")
library(bigreadr)

library(igraph)
library(ggraph)
library(ggrepel)

source("../3_scripts/4_functions.R")
```


Posts: [Coulmont](coulmont.com/blog/2018/12/02/sociologue-reseau-theses-2018/)

[Olivier Gimenez](https://oliviergimenez.github.io/post/network_ecology/)

I am in the data <3

# Get data

```{r load_data}
# Load data
thesis <- read_csv2("../2_clean_data/thesis.csv")
```

We filter the discipline column searching for "evolut|évolut". Sadly, the discipline names can include evolution but still be very large (such as "biological sciences and evolution"). There are also some discipline with the word "evolution" with another meaning: "evolution of terrrestrial systems" for example... So I filter out geology related fields.

We make sure that we only have thesis that were defended, and for which we know the defense date.

```{r thesis_evolution}
# Theses en évolution qui ont été soutenues.
# "Evolut dans discipline : des fois, c'est beaucoup plus large. Should I do it by title?

# 638? That seem not a lot.
thesis_evolution <- thesis %>% 
  filter(str_detect(discipline, 
                    pattern = "Évolut|Evolut|evolut|évolut|Genetique des populations") |
           discipline == "Physiologie et biologie des organismes et populations" | 
           discipline == "Biologie des populations et ecologie") %>% 
  drop_na(date_defense) %>% 
  filter(!str_detect(discipline,
                   pattern = "Terre|planete|lithos|geologie|Geologie|evolutionnaire|Materiaux|materiaux"))
```


We now have `r nrow(thesis_evolution)` PhD, from 1986 to 2018.

We can look at the distribution of the year of publication: I don't know whether the lower number of defended PhDs in the late nineties are because of a large shift towards a different name for the fields.


```{r}
thesis_evolution %>% 
  ggplot(aes(x = date_defense)) +
  geom_histogram() +
  theme_ipsum_rc() +
  labs(title = "Nb of PhD defended in evolution, 1986 - 2018",
       subtitle = "Obviously, some are missing",
       x = "Year",
       y = "Nb of defended PhD",
       caption = "Data: theses.fr")
```

We filter based on the discipline, so the names are *i*) not formatted, and *ii*) change regularly even within doctoral school. It means that we are missing some PhD. 

I tried to search for people I know, and quite a lot of them are in the list, but I found at least one who isn't present as a PhD candidate. 

Since I cannot filter with something as general as "biology" in the discipline, or words like "evolution" or "selection" in the title of the thesis, I will leave it as that.

One could argue that I could try to include PhD with "genetics" in the title, but I decided not to do it, to keep the sample manageable.



# Get comittes

Now that we have the ID of the PhDs in this period related to evolution, we need to get the data about the PhD committee related to these PhDs. To do that, we need to scrap the webpage of each PhD (using the PhD ID) and we use the `rvest` package again.

```{r, eval = FALSE}
thesis_id <- thesis_evolution$thesis_id # get PhD ids

#thesis_id <- head(thesis_id)

total_network <- data.frame(jury_members = "",
                            jury_links = "",
                            these      = "",
                            directeurs = "",
                            advisor_id = "")


for (i in 1:length(thesis_id)) {
  
  # get info on current PhD
  data_phd_evolution <- read_html(paste0("http://www.theses.fr/",
                                      thesis_id[i]) ) 
  
  # get name PhD supervisor for 
  directeurs <- bind_cols(
    directeurs = data_phd_evolution  %>%
      html_nodes("div .donnees-ombre p") %>%
      .[[1]] %>%
      html_nodes("a") %>%
      html_text()
    ,
    advisor_id = data_phd_evolution  %>%
      html_nodes("div .donnees-ombre p") %>%
      .[[1]] %>%
      html_nodes("a") %>%
      html_attr(name="href")
  ) %>% mutate(these = thesis_id[i])
  
  
  # get names of people in commitees
  jury <- bind_cols( 
    jury_members = data_phd_evolution %>%
      html_nodes("div .donnees p a") %>%
      html_text()
    ,
    jury_links = data_phd_evolution %>%
      html_nodes("div .donnees p a") %>%
      html_attr(name="href")
  ) %>% mutate(  these = thesis_id[i] )
  
  # put all together
  network <- jury %>% left_join(directeurs,by="these") 
  total_network <- bind_rows(total_network, network)
}

# Because the process is a bit long, we save the file
save(thesis_evolution,
     total_network,
     file = "../2_clean_data/network_total.RData")
```


# Built network

```{r load_full, echo=FALSE}
load( "../../2_clean_data/network_total.RData")
```


Baptiste Coulmont weighted the different types of links between collegues depending on their role in the PhD process.   
- If two scientists are cosupervisor of a PhD, the link has a weight of 3, because we assume that co-supervising a PhD student creates or is the consequence of a strong link.  

- If one of them is a supervisor and the other is in the PhD commitee, the link has a weight 2, indicating a moderate link.   

- If both researchers are in the same committee, the link is weighted by 1.  


The weight of these links can be added (a lot of people are involved both in co-supervisionsa and committes).


```{r give_weights}
# Link supervisor - supervisor
advisor_advisor <- total_network %>% 
  select(these, directeurs) %>% 
  unique() %>% 
  group_by(these) %>% 
  mutate(N = n()) %>%
  
  filter(N == 2) %>% # keep co-supervision w/ 2 supervisors 
  mutate(rang  = rank(directeurs)) %>% 
  spread(key   = rang, 
         value = directeurs) %>% 
  
  ungroup() %>% 
  select(name_1 = `1`, name_2 = `2`) %>% 
  mutate(poids = 3)


# Link advisor - jury
advisor_jury <- total_network %>% 
  select(name_1 = jury_members,
         name_2 = directeurs) %>% 
  
  filter( name_1 != "") %>%
  mutate(poids = 2) %>%
  group_by(name_1, name_2) %>% 
  # Sum weight over links
  summarize(poids = sum(poids))

# Jury - jury links
jury_jury <- total_network %>% 
  select(jury_members,these) %>% 
  unique() %>% 
  filter(jury_members != "")
```

```{r}
head(advisor_advisor)
head(advisor_jury)
head(jury_jury)
```


Now we use the `graph_from_data_frame()` from the `igraph` package to create the graph.


```{r jury_graph}
# Make non-directed graph for jur_jury
g_j <- graph_from_data_frame(jury_jury,
                              directed = F)

 
# Create the vertex sequence
igraph::V(g_j)$type <- V(g_j)$name %in% jury_jury$jury_members

g_j_1 <- bipartite_projection(g_j, which = "true")


jurys <- as_long_data_frame(g_j_1) %>%
  select(name_1 = `ver[el[, 1], ]`, 
         name_2 = `ver2[el[, 2], ]`, 
         poids  = weight)

reseau_petit <- bind_rows(advisor_advisor,
                          advisor_jury,
                          jurys) %>%
  group_by(name_1, name_2) %>% 
  summarize(poids = sum(poids)) # data.frame from which the network will be created
```

# Plot

We correlate the size of the point to a *betweenness* measure between nodes.The width of edges correlates with the summed weight of the link between the two scientists (the strenght of their connection).
We determine communities trough a [cluster_walktrap]("https://igraph.org/r/doc/cluster_walktrap.html") algorithm, but to be honest, it's mainly to add colour to the graph. The `cluster_walktrap()` function, from the `igraph` package tries to find densely connected subgraphs (communities) via random walks. The idea is that short random walks tend to stay in the same community. 


## Plot the full network (1986 - 2018)


```{r make_network}
g <- graph_from_data_frame(reseau_petit, 
                           directed = F) 

g <- simplify(g, edge.attr.comb = sum)

V(g)$degres <- degree(g)

# Get surname only
V(g)$label  <- gsub("^\\S+\\s+(.+)$","\\1",V(g)$name)


# determine communities
# step = the length of the random walk to perform
V(g)$communaute <- as.character(cluster_walktrap(g,
                                                 steps = 10)$membership)  # 15 originellement

V(g)$closeness <- (5*closeness(g))^10

# network metric betweeness
V(g)$btwns       <- betweenness(g) 
V(g)$eigen_centr <- eigen_centrality(g)$vector


# delete edges with weight <= 4
g <- delete_edges(g, which(E(g)$poids < 4)) # 5 initiallement



# to which community you belong
V(g)$cluster_number <- clusters(g)$membership 

g <- induced_subgraph(g,
                      V(g)$cluster_number == which( max(clusters(g)$csize) == clusters(g)$csize) )

# width of edge proportional to weight
E(g)$weight <- 1/E(g)$poids 


# do not display all names
V(g)$label <- ifelse(V(g)$degres < 8, 
                     "", 
                     V(g)$label)  # 20 initialement
```



```{r plot}
graphe_1 <- ggraph(g,
                   layout = "igraph",
                   algorithm = "fr") + 
  
  geom_edge_link(aes(width = 0.1*poids), alpha = 0.1, 
                 end_cap   = circle(5, 'mm'), 
                 start_cap = circle(5, 'mm')) +
  
  geom_node_point(aes(size = eigen_centr),
                  color = "white", alpha = 1) +
  
  geom_node_point(aes(color = communaute,
                      size = eigen_centr), 
                  alpha = 0.5) +
  
  scale_size_area(max_size = 20) +
  
  geom_node_text(aes(label = label),
                 size = 3, 
                 repel = T,
                 box.padding = 0.15) +
  
  labs(title = "Réseaux des évolutionistes",
       subtitle="Soutenances de thèses entre 2015 et 2018",
       caption="Sources : theses.fr \n Code par B. Coulmont, modifié par O. Gimenez") +
  
  theme_graph(foreground = 'white', fg_text_colour = 'white') +
  
  theme(legend.position = "none",
        text = element_text(size = 16),
        plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "line"))

# ggsave("network1.png",
#        graphe_1,
#        units = "cm",
#        width = 30,
#        height = 25)


# png(filename = "network1.png",
#      units = "cm",
#        width = 30,
#        height = 25)
# graphe_1
# dev.off()

```


That's a lot to digest.

The first thing that I notice is internal: it is a particularly good feeling when you know the people that are on the graph.

The second thing that I notice was a lack of connections between my two PhD supervisors. I dug a bit and realized, several years after defending my PhD that the data on the theses.fr is wrong, only one of my former advisor is listed. I asked for the correction, of course.

I wish I could plot more names, but my computer crashes if I try to. Every. Single. Time. So we will have to do with this level of precision, at least with the full dataset.

I can recognize groups of people whom I know work together, publish together, and have a scientific connection, which is quite reassuring. I am surprised that the spatial closeness is not always a good predictor of network: I was expecting people of Montpellier to be more close. But they are scattered all over the place. It means that we have many subfield represented here, and that scientists have a wide network in other labs in general.


## Plot the network 2015 to 2018

Because the dataset was so large, it is difficult to plot more names. So let's see what the network looks like in the past three years.

I wrotte a couple of functions to perform the same scrapping as above and save some space. See **LINK** for the source file

```{r source_functions}
source("../3_scripts/4_functions.R")
```

```{r}
thesis_evolution_2015_2018 <- thesis_evolution %>% 
  filter(year_defense > 2014) 
```

We know have a data.frame with `r nrow(thesis_evolution_2015_2018)` lines.


```{r scrap2015_2018, eval=FALSE}
scrapped_2015_2018 <- scrap_phd_webpages(thesis_evolution_2015_2018)

save(thesis_evolution_2015_2018,
     scrapped_2015_2018,
     file = "../../2_clean_data/scrapped_2015_2018.RData")
```


```{r load_2015, echo=FALSE}
load( "../../2_clean_data/scrapped_2015_2018.RData")
```



```{r plot_2015_2018}
graphe_2 <- make_network_full(scrapped_2015_2018,
                  my_waltrap = 10,
                  my_edge    = 3,
                  my_degree  = 4,
                  my_title   = "Network of evolutionary biologists",
                  my_subtitle = "Based on French PhD advising and committes, 2015 - 2018")




# png(filename = "network2.png", res = 100)
#      #units = "cm", width = 30, height = 25, res = 300)
# graphe_2
# dev.off()
```





## Plot the network in Montpellier

Since I did my PhD in Montpellier, I was very curious to see the local network.


```{r}
thesis_evolution_Montpellier <- filter(thesis_evolution,
                             str_detect(university,
                                        pattern = "Montpellier"))
```


```{r scrap_montpel, eval=FALSE}
scrapped_Montpellier <- scrap_phd_webpages(thesis_evolution_Montpellier)

save(thesis_evolution_Montpellier,
     scrapped_Montpellier,
     file = "../../2_clean_data/scrapped_Montpellier.RData")
```

```{r load_montpel, echo = FALSE}
load( "../../2_clean_data/scrapped_Montpellier.RData")
```


```{r plot_montpell}
graphe_3 <- make_network_full(scrapped_Montpellier,
                  my_waltrap = 7,
                  my_edge    = 1,
                  my_degree  = 2,
                  my_title   = "Network of evolutionary biologists",
                  my_subtitle = "Based on French PhD advising and committes, Montpellier")


# png(filename = "network3.png")
#      #units = "cm", width = 30, height = 25, res = 300)
# graphe_3
# dev.off()
# 
# ggsave("network3.png",
#        graphe_3,
#        units = "cm",
#        width = 30,
#        height = 25)
```


